---
title: "Report on Prison Admission Trends in the U.S."
author: "Simon Cao"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
source("../source/analysis.R")
```

## Introduction
<p>The data set used as the basis of the analysis in this report is from [this repository](https://github.com/vera-institute/incarceration-trends/blob/master/incarceration_trends.csv).
This data set from the Vera Institute of Justice provides county-level data on
the topic of incarceration in the U.S. with `r format(obs_num,big.mark =",")`
observations and `r format(var_num,big.mark =",")` variables. </p>
<p>The United States has a significant social problem in racial justice, 
especially in racial discrimination by law enforcement. In this analysis, 
we focus on the U.S. prison admission rate of people of color to identify the 
racial discrimination involved. </p>

# Summary Information
<p>In general, people of color suffer significant larger prison admission rate
than average with the black population stands out. For example, In 2016, the
average prison admission rate is
`r format(avg_prison_admission_rate,big.mark =",")`, but for the
 black population, the number is significantly higher, with an average of
`r format(avg_black_prison_adm_rate,big.mark =",")`, and a maximum of
`r format(max_black_prison_adm_rate,big.mark =",")`, comparing to the white
population average `r format(avg_white_prison_adm_rate,big.mark =",")`.
However, through out the years, this condition seems to be improving. 
From 2011 to 2016, the average prison admission rate for black people decreased
by `r format(black_prison_adm_rate_diff,big.mark =",")`. All detailed numbers
are listed below.
 </p>
```{r, echo=FALSE}
kable(summary_table)
```
- The prison admission rate is established on a per 100,000 people basis.

## Trends over time plot
<p>The following time trend plot gives a sense of the difference in prison
admission between race groups changes of the situation over time.
From the graph, we could see that the black population has significantly higher
prison admission rate than other race groups. However, it seems like the
situation is improving over the years.</p>
```{r, echo=FALSE}
time_trend_plot
```


## Variable Comparison Plot
<p>The plot compared the prison admission rate with the black population
for all 50 states in 2016. By fitting linear regression to the scatter plot,
we find that with more black population, prison admission actually drops rather
than stays on the same level. This could imply states with more black population
may have fairer law enforcement based on more influence.</p>
```{r, echo=FALSE, fig.width = 8, fig.asp = .8}
suppressWarnings(print(variable_comparison_plot))
```

## Map plot
<p>The following is a heat map of prison admission rate of black people in 2016.
The map reveals geological distribution of discrimination, high lighting more
severe racial injustice in central America with Missouri having the highest
prison admission rate for black people. </p>
```{r, echo=FALSE,fig.asp = .55}
map_plot
```