---
title: "Report on Prison Admission Trends in the U.S."
author: "Simon Cao"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
source("../source/analysis.R")
```

## Introduction
<p>The data set used in this analysis is from [this repository](https://github.com/vera-institute/incarceration-trends/blob/master/incarceration_trends.csv).
The dataset provides county-level data on incarceration in the 
U.S. with `r format(obs_num,big.mark =",")` observations and
`r format(var_num,big.mark =",")` variables. </p>
<p>The United States has a significant problem in racial justice, especially in 
racial discrimination by law enforcement. In this analysis, we focus on the 
U.S. prison admission rate of people of color to identify the unfairness. </p>

# Summary Information
- The prison admission rate below is established on a per 100,000 people basis.
It is clear that through 2011 to 2016, people of color suffer significant larger
 prison admission rate than average with the black population stands out (
greater than 1000 admission rate).
Detailed numbers are listed below.

```{r, echo=FALSE}
kable(summary_table)
```

## Trends over time plot 

```{r, echo=FALSE}
time_trend_plot
```


## Variable Comparison Plot
```{r, echo=FALSE, fig.width = 8, fig.asp = .8}
suppressWarnings(print(variable_comparison_plot))
```

## Map plot
```{r, echo=FALSE,fig.asp = .55}
map_plot
```